Preparing your manuscript

### 2025-05-26 article/article.qmd
Add SST gradient

file_list <- c(Sys.glob(here("data/eke/*hist*mean.nc")), 
               Sys.glob(here("data/deltat/*hist*mean.nc")),
               Sys.glob(here("data/tos/*hist*mean.nc")))

data <- purrr::map(file_list, function(f) {

@@

tas <- purrr::map(Sys.glob(here("data/tas/*hist*.nc")), function(f) {

  message(f)
  meta <- unglue::unglue(basename(f), patterns = c("{var}_day_{model}_{scenario}_{member}_{grid}_20150101-21001231.nc"))

  file_scenario <- Sys.glob(here(paste0("data/", meta[[1]][["var"]], "/", 
                                        meta[[1]][["var"]], "_day_", meta[[1]][["model"]], "_ssp585*")))

@@

  ReadNetCDF(f, vars = c(hist = var)) |> 
    _[, let(ssp585 = ReadNetCDF(file_scenario, vars = c(ssp585 = var), out = "vector")[[1]])] |> 
    _[, let(model = meta[[1]][["model"]],
            var = "dt_dy",
            lat = round(lat, digits = 8),
            hist = metR:::.derv(hist, lat)*(180/pi/6371000),
            ssp585 = metR:::.derv(ssp585, lat)*(180/pi/6371000)), by = .(lon)] |> 
    _[, let(change = ssp585 / hist,
            diff = ssp585 - hist)]

@@

data <- rbind(data, tas) |> 
  _[, let(land = MaskLand(lon, lat))] |> 
  _[, let(land = factor(land, labels = c("Sea", "Land")))]

>>>>

file_list <- c(Sys.glob(here("data/eke/*hist*mean.nc")), 
               Sys.glob(here("data/deltat/*hist*mean.nc")))

delta <- purrr::map(file_list, function(f) {

@@

tas <- purrr::map(Sys.glob(here("data/tas/*hist*mean.nc")), function(f) {

  message(f)
  meta <- unglue::unglue(basename(f), patterns = c("{var}_{model}_historical_mean.nc"))

  file_scenario <- Sys.glob(here(paste0("data/", meta[[1]][["var"]], "/", 
                                        meta[[1]][["var"]], "_", meta[[1]][["model"]], "_ssp585*")))

@@

  ReadNetCDF(f, vars = c(hist = var)) |> 
    _[, let(ssp585 = ReadNetCDF(file_scenario, vars = c(ssp585 = var), out = "vector")[[1]])] |> 
    _[, let(model = meta[[1]][["model"]],
            var = var,
            lat = round(lat, digits = 8),
            hist = metR:::.derv(hist, lat)*(180/pi/6371000),
            ssp585 = metR:::.derv(ssp585, lat)*(180/pi/6371000)), by = .(lon)] |> 
    _[, let(change = ssp585 / hist,
            diff = ssp585 - hist)]

@@

tos <- purrr::map(Sys.glob(here("data/tos/*hist*mean.nc")), function(f) {

  message(f)
  meta <- unglue::unglue(basename(f), patterns = c("{var}_{model}_historical_mean.nc"))

  file_scenario <- Sys.glob(here(paste0("data/", meta[[1]][["var"]], "/", 
                                        meta[[1]][["var"]], "_", meta[[1]][["model"]], "_ssp585*")))

  ReadNetCDF(f, vars = c(hist = var)) |> 
    _[, let(ssp585 = ReadNetCDF(file_scenario, vars = c(ssp585 = var), out = "vector")[[1]])] |> 
    _[, let(model = meta[[1]][["model"]],
            lat = round(lat, digits = 8),
            var = var,
            dt_dy_hist = metR:::.derv(hist, lat)*(180/pi/6371000),
            dt_dy_ssp585 = metR:::.derv(ssp585, lat)*(180/pi/6371000)), by = .(lon)] |> 
    _[, let(cos_lat = cos(lat*180/pi))] |> 
    _[, let(dt_dx_hist = metR:::.derv(hist, lon)*(180/pi/6371000)*cos_lat,
            dt_dx_ssp585 = metR:::.derv(ssp585, lon)*(180/pi/6371000)*cos_lat), by = .(lat)] |> 
    _[, let(change = sqrt(dt_dx_ssp585^2 + dt_dy_ssp585^2) / sqrt(dt_dx_hist^2 + dt_dy_hist^2),
            diff = sqrt(dt_dx_ssp585^2 + dt_dy_ssp585^2) - sqrt(dt_dx_hist^2 + dt_dy_hist^2))] |> 
    _[, .(time, lat, lon, hist, ssp585, model, var, change, diff)]

mask <- tos[var == "tos" & model == "NorESM2-MM", .(lon, lat, hist)] |> 
  _[, let(land = fifelse(!is.na(hist), "Land", "Sea"))] |> 
  _[, .(lon, lat, land)]

data <- rbindlist(list(delta, tas, tos)) |> 
  mask[x = _, on = .NATURAL]
  # _[, let(land = MaskLand(lon, lat))] |> 
  # _[, let(land = factor(land, labels = c("Sea", "Land")))]
